---
title: "ManyFaces validation/norming ratings"
subtitle: "Compiling ratings by face"
format: 
  html:
    code-fold: true
    df-print: kable
    embed-resources: true
    fig-height: 5
    fig-width: 8
    toc: true
execute: 
  error: true
---

```{r}
#| message: false
library(tidyverse)

# Load custom functions
source("data/custom_functions.R")
```

## Load data
```{r}
# Rating data
data_exp <- read_csv("data/manyfaces_ratings_exp_cleaned.csv",
                     show_col_types = FALSE)

exp_levels_neutral <- c("attractive", "trustworthy", "dominant", 
                        "memorable", "gender-typical")

exp_levels_neutral_unstd <- c("attractive (unstd)", "trustworthy (unstd)",
                              "dominant (unstd)")

exp_levels_emo_cat <- c("emotion 1", "emotion 2", "emotion 3", 
                        "emotion 4", "emotion 5", "emotion 6")

exp_levels_emo_int <- c("fear", "anger", "disgust", 
                        "happiness", "sadness", "surprise")

# Questionnaire data
data_qu <- read_csv("data/manyfaces_ratings_quest_cleaned.csv",
                    show_col_types = FALSE)

# Model data (available via UK Data Service ReShare)
data_model <- read_csv("ReShare_data/manyfaces_model_data_cleaned.csv",
                       show_col_types = FALSE)

rainbow <- c("firebrick", "darkorange", "goldenrod",
             "darkgreen", "dodgerblue3", "darkorchid4")
```


## Standardised neutral trait ratings

```{r}
std_neutral <- data_exp |>
  filter(exp %in% exp_levels_neutral) |>
  mutate(dv = as.integer(dv),
         exp = factor(exp, exp_levels_neutral)) |> 
  separate_wider_delim(trial_name,
                       delim = "_",
                       names = c("lab_id", "model_id"),
                       too_many = "drop")
```

```{r}
std_neutral |>
  summarise(N = n(),
            mean_rating = mean(dv),
            .by = c("exp", "lab_id", "model_id")) |> 
  summarise(N = first(N),
            M = mean(mean_rating),
            SD = sd(mean_rating),
            .by = "exp") |> 
mutate(across(where(is.numeric) & !any_of("N"), ~ round(.x, 2)))
```

```{r fig.height=5, fig.width=8}
std_neutral |>
  ggplot(aes(x = dv, y = after_stat(count / sum(count)), fill = exp)) +
  geom_histogram(binwidth = 1, show.legend = FALSE, color = "black") +
  facet_wrap(~exp, ncol = 3, axes = "all_x", drop = FALSE) +
  labs(title = "Standardized neutral trait ratings",
       x = "Rating",
       y="Proportion") +
  scale_x_continuous(breaks = 1:7) +
  scale_fill_manual(values = rainbow, drop = FALSE) +
  theme_minimal() +
  theme(strip.text = element_text(size = 12, face = "bold"),
        axis.text = element_text(size = 10),
        axis.title = element_text(size = 12, face = "plain"),
        title = element_text(size = 13, face = "bold"))
```



Save mean rating for each face for each trait
```{r}
std_neutral_means <- std_neutral |> 
  summarise(mean_rating = mean(dv),
            rater_N = n(),
            .by = c("exp", "lab_id", "model_id")) |> 
  pivot_wider(id_cols = c(lab_id, model_id),
              names_from = exp,
              #values_from = c(mean_rating, rater_N))
              values_from = mean_rating)
```

#### Attractiveness

Please note: The heatmaps visualise the frequency distribution of ratings for each model (y-axis) from each lab (shown in separate panels). Darker tiles indicate that more raters gave a particular rating (x-axis) to that model.

```{r}
#| fig-height: 10
#| fig-width: 10

heatmap(std_neutral, "attractive", "Attractiveness")
```

#### Trustworthiness

```{r}
#| fig-height: 10
#| fig-width: 10

heatmap(std_neutral, "trustworthy", "Trustworthiness")
```

#### Dominance

```{r}
#| fig-height: 10
#| fig-width: 10

heatmap(std_neutral, "dominant", "Dominance")
```

#### Memorability

```{r}
#| fig-height: 10
#| fig-width: 10

heatmap(std_neutral, "memorable", "Memorability")
```

#### Gender typicality

```{r}
#| fig-height: 10
#| fig-width: 10

heatmap(std_neutral, "gender-typical", "Gender Typicality")
```

##### Perceived typicality vs self-reported gender

```{r}
gender <- std_neutral |> 
  filter(exp == "gender-typical") |> 
  left_join((data_model |> dplyr::select(lab_id, model_id, gender_rec)), by = c("lab_id", "model_id"))
```

```{r}
gender |> 
  ggplot(aes(x = dv, fill = gender_rec)) +
  geom_histogram(binwidth = 1) +
  scale_fill_manual(values = c("hotpink", "lightblue", "orchid", "gray")) +
  labs(x = "Gender typicality",
       fill = "Self-reported gender")
```

## Unstandardised neutral trait ratings

```{r}
unstd_neutral <- data_exp |>
  filter(exp %in% exp_levels_neutral_unstd) |>
  mutate(dv = as.integer(dv),
         exp = factor(exp, exp_levels_neutral_unstd)) |> 
  separate_wider_delim(trial_name,
                       delim = "_",
                       names = c("lab_id", "model_id"),
                       too_many = "drop")
```

```{r}
unstd_neutral |>
  summarise(N = n(),
            mean_rating = mean(dv),
            .by = c("exp", "lab_id", "model_id")) |> 
  summarise(N = first(N),
            M = mean(mean_rating),
            SD = sd(mean_rating),
            .by = "exp") |> 
mutate(across(where(is.numeric) & !any_of("N"), ~ round(.x, 2)))
```

```{r fig.height = 2.75, fig.width=8}
unstd_neutral |>
  ggplot(aes(x = dv, y = after_stat(count / sum(count)), fill = exp)) +
  geom_histogram(binwidth = 1, show.legend = FALSE, color = "black") +
  facet_wrap(~exp, ncol = 5, drop = FALSE) +
  labs(title = "Unstandardized neutral trait ratings",
       x = "Rating",
       y = "Proportion") +
  scale_x_continuous(breaks = 1:7) +
  scale_fill_manual(values = rainbow, drop = FALSE) + 
  theme_minimal() +
  theme(strip.text = element_text(size = 12, face = "bold"),
        axis.text = element_text(size = 10),
        axis.title = element_text(size = 12, face = "plain"),
        title = element_text(size = 13, face = "bold"))
```

Save mean rating for each face for each trait
```{r}
unstd_neutral_means <- unstd_neutral |> 
  summarise(mean_rating = mean(dv),
            rater_N = n(),
            .by = c("exp", "lab_id", "model_id")) |> 
  pivot_wider(id_cols = c(lab_id, model_id),
              names_from = exp,
              #values_from = c(mean_rating, rater_N))
              values_from = mean_rating)
```

#### Attractiveness (Unstandardised)

```{r}
#| fig-height: 10
#| fig-width: 10

heatmap(unstd_neutral, "attractive (unstd)", "Attractiveness (Unstandardised)")
```

#### Trustworthiness (Unstandardised)

```{r}
#| fig-height: 10
#| fig-width: 10

heatmap(unstd_neutral, "trustworthy (unstd)", "Trustworthiness (Unstandardised)")
```

#### Dominance (Unstandardised)

```{r}
#| fig-height: 10
#| fig-width: 10

heatmap(unstd_neutral, "dominant (unstd)", "Dominance (Unstandardised)")
```

## Emotion categorisations

```{r}
emo_cat <- data_exp |>
  filter(exp %in% exp_levels_emo_cat) |>
  mutate(exp = factor(exp, exp_levels_emo_cat),
         dv = factor(dv, levels = c("anger", "disgust", "fear", 
                                    "happiness", "sadness", "surprise",
                                    "other"))) |> 
  separate_wider_delim(trial_name,
                       delim = "_",
                       names = c("lab_id", "model_id", NA, "model_expression", NA),
                       too_many = "drop") |> 
  mutate(model_expression = forcats::fct_recode(model_expression,
                                                anger = "ang",
                                                disgust = "dis",
                                                fear = "fea",
                                                happiness = "hap",
                                                sadness = "sad",
                                                surprise = "sur"))
```

```{r}
emo_cat |> 
  group_by(model_expression) |> 
  count(model_expression, dv) |> 
  mutate(prop = n / sum(n)) |>
  ungroup() |> 
  dplyr::select(-n) |> 
  pivot_wider(names_from = dv,
              values_from = prop) |> 
  relocate(other, .after = surprise) |> 
  mutate(across(where(is.numeric), ~ round(.x, 2)))
```

```{r fig.height=5, fig.width=8}
emo_levels <- levels(emo_cat$model_expression)
emo_labels <- paste(emo_levels, "faces")

emo_prop <- emo_cat |> 
  group_by(model_expression) |> 
  count(model_expression, dv) |> 
  mutate(prop = n / sum(n)) |>
  ungroup() 

emo_prop |>
  mutate(emo = factor(model_expression, emo_levels, emo_labels)) |>
  ggplot(aes(x = dv, y=prop, fill = dv)) +
  geom_point(aes(x = x, colour = I(fill), fill = I(fill)), 
             data.frame(emo = factor(emo_levels, emo_levels, emo_labels),
                        x = 1:6,
                        fill = rainbow),
             size = 6.5, y = -60, shape = 18, show.legend = FALSE) +
  geom_bar(color = "transparent", stat="identity") +
  facet_wrap(~emo, axes = "all_x", drop = FALSE) +
  scale_x_discrete(labels = c("", "", "", "", "", "", "")) +
  scale_fill_manual(values = c(rainbow, "grey"), drop = FALSE) +
  labs(title = "Standardized emotion categorizations",
       x = "Categorization",
       y = "Proportion",
       fill = "Categorized as") +
  coord_cartesian(clip = "off") +
  theme_minimal() +
  theme(axis.ticks.x = element_blank(),
        strip.text = element_text(size = 12, face = "bold"),
        axis.text = element_text(size = 10),
        axis.title = element_text(size = 12, face = "plain"),
        title = element_text(size = 13, face = "bold"))

# this shows counts rather than proportions, but includes downward arrows for most-chosen category

#emo_cat |>
#  mutate(emo = factor(model_expression, emo_levels, emo_labels)) |>
#  ggplot(aes(x = dv, fill = dv)) +
#  geom_point(aes(x = x, colour = I(fill), fill = I(fill)), 
#             data.frame(emo = factor(emo_levels, emo_levels, emo_labels),
#                        x = 1:6,
#                        fill = rainbow),
#             size = 6.5, y = -60, shape = 18, show.legend = FALSE) +
#  geom_bar(color = "transparent") +
#  facet_wrap(~emo, axes = "all_x", drop = FALSE) +
#  scale_x_discrete(labels = c("", "", "", "", "", "", "")) +
#  scale_fill_manual(values = c(rainbow, "grey"), drop = FALSE) +
#  labs(title = "Standardized emotion categorizations",
#       x = "",
#       fill = "Categorized as") +
# coord_cartesian(clip = "off") +
#  theme(axis.ticks.x = element_blank())
```

Save proportion of raters choosing each emotion for each face
```{r}
emo_cat_props <- emo_cat |> 
  group_by(lab_id, model_id, model_expression) |> 
  count(model_expression, dv) |> 
  group_by(lab_id, model_id, model_expression) |>
  mutate(prop = n / sum(n)) |>
  ungroup() |> 
  dplyr::select(-n) |> 
  pivot_wider(names_from = dv,
              values_from = prop) |> 
  relocate(other, .after = surprise)
```

## Emotion intensity ratings

```{r}
emo_int <- data_exp |>
  filter(exp %in% exp_levels_emo_int) |>
  mutate(dv = as.integer(dv),
         exp = factor(exp, exp_levels_emo_int)) |> 
  separate_wider_delim(trial_name,
                       delim = "_",
                       names = c("lab_id", "model_id"),
                       too_many = "drop")
```

```{r}
emo_int |>
  summarise(N = n(),
            mean_rating = mean(dv),
            .by = c("exp", "lab_id", "model_id")) |> 
  summarise(N = first(N),
            M = mean(mean_rating),
            SD = sd(mean_rating),
            .by = "exp") |> 
mutate(across(where(is.numeric) & !any_of("N"), ~ round(.x, 2)))
```

```{r fig.height=5, fig.width=8}
emo_int |>
  ggplot(aes(x = dv, y = after_stat(count / sum(count)), fill = exp)) +
  geom_histogram(binwidth = 1, show.legend = FALSE, color = "black") +
  facet_wrap(~exp, ncol = 3, axes = "all_x", drop = FALSE) +
  labs(title = "Standardized emotion intensity ratings",
       x = "Rating",
       y = "Proportion") +
  scale_fill_manual(values = rainbow, drop = FALSE) +
  scale_x_continuous(breaks = 1:7) + 
  theme_minimal() +
  theme(strip.text = element_text(size = 12, face = "bold"),
        axis.text = element_text(size = 10),
        axis.title = element_text(size = 12, face = "plain"),
        title = element_text(size = 13, face = "bold"))
```

Save mean rating for each face for each trait
```{r}
emo_int_means <- emo_int |> 
  summarise(mean_rating = mean(dv),
            rater_N = n(),
            .by = c("exp", "lab_id", "model_id")) |> 
  rename(model_expression = exp,
         mean_intensity_rating = mean_rating) |> 
  dplyr::select(-rater_N)
```

## Age Ratings

```{r}
dv_levels <- seq(20, 85, 5)
dv_labels <- paste(dv_levels-4, "-", dv_levels )
dv_labels[14] <- "81+"

age <- data_exp |>
  filter(exp == "age") |>
  mutate(dv = factor(dv, dv_levels, dv_labels)) |> 
  separate_wider_delim(trial_name,
                       delim = "_",
                       names = c("lab_id", "model_id"),
                       too_many = "drop")
```

```{r}
age |>
  ggplot(aes(x = dv)) +
  geom_bar(color = "black", fill = "gray") +
  scale_x_discrete(drop = FALSE) +
  labs(title = "Age Ratings",
       x = "") +
  theme_minimal()
```

Save modal age for each face
```{r}
age_modes <- age |>
  group_by(lab_id, model_id) |>
  summarise(out = list(get_mode_info(dv)),
            .groups = "drop") |>
  unnest_wider(out)
```

```{r}
#| fig-height: 10
#| fig-width: 10

age |>
  count(lab_id, model_id, dv) |>
  ggplot(aes(x = dv, y = model_id, fill = n)) +
  geom_tile() +
  facet_wrap(~lab_id) +
  scale_fill_viridis_c() +
  labs(x = "Age", y = NULL) +
  theme(legend.position = "none", 
        axis.text.x = element_text(angle = 90))
  
```

### Actual vs perceived age

Relate actual age to perceived modal age for each face
```{r}
age_perc <- age_modes |> 
  mutate(age_mid = age_mode) |> 
  separate(age_mid, into = c("age_start", "age_end"), sep = "-", convert = TRUE) |> 
  mutate(age_perc = (age_start + age_end) / 2) |> 
  select(-age_start, -age_end)

age_by_face <- data_model |> 
  dplyr::select(lab_id, model_id, age) |> 
  rename(age_actual = age) |> 
  left_join(age_perc, by = c("lab_id", "model_id"))
```

```{r}
# Get stats
age_by_face_cortest <- cor.test(age_by_face$age_actual, age_by_face$age_perc)
age_by_face_r <- age_by_face_cortest$estimate
age_by_face_p <- age_by_face_cortest$p.value
label_text <- sprintf("r = %.2f, p = %.3f", age_by_face_r, age_by_face_p)
```

```{r}
age_by_face |> 
  filter(!is.na(age_actual) & !is.na(age_perc)) |> 
  ggplot(aes(x = age_actual, y = age_perc)) +
  geom_point() +
  geom_smooth(method = "lm",  formula = y ~ x, se = TRUE, colour = "black") +
  annotate("text", x = 65, y = 20, label = label_text, hjust = 0) +
  labs(x = "Actual Age", y = "Perceived Age [Mode]") +
  theme_bw()
```


## Compile ratings
```{r}
ratings_traits <- std_neutral_means |> 
  full_join(unstd_neutral_means, by = c("lab_id", "model_id"))

ratings_emotions <- emo_cat_props |> 
  left_join(emo_int_means, by = c("lab_id", "model_id", "model_expression"))
```

```{r}
write_csv(ratings_traits, "results/manyfaces_ratings_traits.csv")
write_csv(ratings_emotions, "results/manyfaces_ratings_emotions.csv")
```


